---
title: "Hoppers Data"
author: "Nicholas Heimbuch"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, echo=FALSE, warning=FALSE, message = FALSE, results='hide'}
knitr::opts_chunk$set(echo = FALSE, message = FALSE)
pacman::p_load(tidyverse, readxl)
```


```{r data}
trt <- read_xlsx("../data/HopperzR.xlsx",
                       "Treatment")
#biomass data
fb <- read_xlsx("../data/HopperzR.xlsx",
               "Forage Biomass") %>%
        full_join( trt ,
                   by="Plot")

#Abundance data
ha<- read_xlsx("../data/HopperzR.xlsx",
               "Hopper Abundance")%>%
      full_join(trt,
                by="Plot")

#Forage quality data
cn<- read_xlsx("../data/HopperzR.xlsx",
          "CN Analysis") %>%
      full_join( trt,
                 by="Plot")

#Density per square meter
ha <- read_xlsx("../data/HopperzR.xlsx",
               "Hopper Abundance") %>%
        full_join(trt,
                  by="Plot")
ring = 2.5 #sum area of five rings (square meters)
```


## Biomass

We predicted that the difference in biomass would decrease with time since fire. 

```{r}
fb %>%
  mutate(atw= WWWeight/WWT,
         atw= ifelse(Treatment=="Summer", Totalweight, atw))%>%
  select(Treatment, Plot, Trial, atw)%>%
  separate(Trial, into= c("Trt", "rep"), sep = " ")%>%
  mutate(rep = case_when(
                rep== "1"~"A", 
                rep== "2"~"B",
                TRUE~ rep),
        Treatment = case_when(
          Treatment== "Summer"~"Unburned",
          TRUE~ Treatment),
    days= ifelse(rep== "A", 40, 33) )%>%
  pivot_wider(names_from = Trt, 
              values_from = atw)%>%
  mutate(diff= (Control-Exclosure)/(Control)*100, 
         rate= diff/days)%>%
group_by(Treatment) %>%
  summarize(Meandiff= mean(rate),
          SEdiff= sd(rate)/sqrt(n())) %>%
  ggplot()+
  geom_hline(yintercept = 0, lty=3)+
  geom_errorbar(aes(x=Treatment,
                    ymin=Meandiff-SEdiff,
                    ymax=Meandiff+SEdiff
                    ),
                position = position_dodge(width=0.2),
                width= 0.2)+
  geom_point(aes(x=Treatment,
                 y=Meandiff
                 ),
             position = position_dodge(width=0.2))+
  labs(x= "Season of Last Burn",
       y= "Daily Rate of \n Grasshopper Removal (%)")
              
```

## Biomass Loss Per Hectare

Burned areas saw heavy grasshopper removal

```{r}
fb %>%
  filter(Treatment != "Summer")%>%
  select(Treatment, Plot, Trial, Totalweight)%>%
  separate(Trial, into= c("Trt", "rep"), sep = " ")%>%
  mutate(rep = case_when(
    rep== "1"~"A", 
    rep== "2"~"B",
    TRUE~ rep),
    days= ifelse(rep== "A", 40, 33) )%>%
  pivot_wider(names_from = Trt, 
              values_from = Totalweight)%>%
  #exclosure is half meter squared
    mutate(gdiff1m= ((Control-Exclosure)*2),
         kgdiffha= gdiff1m*10,
           rate= kgdiffha/(days))%>%
  group_by(Treatment) %>%
  summarize(Meandiff= mean(rate),
            SEdiff= sd(rate)/sqrt(n())) %>%
  mutate(Meandiff= -Meandiff)%>%
  ggplot()+ theme_bw()+
  geom_errorbar(aes(x=Treatment,
                    ymin=Meandiff-SEdiff,
                    ymax=Meandiff+SEdiff
  ),
  position = position_dodge(width=0.2),
  width= 0.2)+
  geom_point(aes(x=Treatment,
                 y=Meandiff
  ),
  position = position_dodge(width=0.2))+
  labs(x= "Season of Last Burn",
       y= "Rate of Removal \n (kg/ha/day)")
```

## Abundance

we predicted that abundance would decrease with time since fire. 

```{r abundance graph}
ha %>%
  mutate(#Date= format(Date, "%d-%B"),
         Hoppers= as.numeric(trimws(Hoppers)) )%>%
  mutate(Treatment = case_when(
    Treatment== "Summer"~"Unburned",
    TRUE~ Treatment))%>%
  group_by(Treatment, Date)%>%
    summarize(MeanHoppers= mean(Hoppers, na.rm=T),
    SEHoppers= sd(Hoppers)/sqrt(n()))%>%
 
ggplot()+
  geom_line(aes(x=Date, y=MeanHoppers,
                color=Treatment))+
  geom_errorbar(aes(x=Date,
                    ymin=MeanHoppers-SEHoppers,
                    ymax=MeanHoppers+SEHoppers,
                    color=Treatment),
  position = position_dodge(width=5),
  width= 5)+
  geom_point(aes(x=Date,
                 y=MeanHoppers,
                 color=Treatment),
  position = position_dodge(width=5))+
  labs(x= "Date",
       y= "Abundance")
```

We found that fall and spring burn treatments were higher in abundance than unburned treatments. 


## Density per Hectare

The densities climbed in burned plots but remained consistent in unburned plots. 

```{r}
ha %>%
  mutate(#Date= format(Date, "%d-%B"),
    Hoppers= as.numeric(trimws(Hoppers)) )%>%
  mutate(Treatment = case_when(
    Treatment== "Summer"~"Unburned",
    TRUE~ Treatment),
    Hoppers= Hoppers/ring)%>%
  group_by(Treatment, Date)%>%
  summarize(MeanHoppers= mean(Hoppers, na.rm=T),
            SEHoppers= sd(Hoppers)/sqrt(n()))%>%
  
  ggplot()+ theme_bw()+
  geom_line(aes(x=Date, y=MeanHoppers,
                color=Treatment))+
  geom_errorbar(aes(x=Date,
                    ymin=MeanHoppers-SEHoppers,
                    ymax=MeanHoppers+SEHoppers,
                    color=Treatment),
                position = position_dodge(width=5),
                width= 5)+
  geom_point(aes(x=Date,
                 y=MeanHoppers,
                 color=Treatment),
             position = position_dodge(width=5))+
  labs(x= "Date",
       y= "Density (Grasshoppers/m^2)")
```

## Forage Quality

We predicted forage quality to be highest in the most recently burned plots. 

```{r crude protein graph}
cn %>% 
group_by(Treatment, Organ)%>%
  summarize(MeanProtein= mean(Protein),
            SEProtein= sd(Protein)/sqrt(n()))%>%
  mutate(Treatment = case_when(
    Treatment== "Summer"~"Unburned",
    TRUE~ Treatment))%>%
  ggplot()+
  geom_errorbar(aes(x=Treatment,
                    ymin=MeanProtein-SEProtein,
                    ymax=MeanProtein+SEProtein,
                    color=Organ),
                position = position_dodge(width=0.2),
                width= 0.2)+
  geom_point(aes(x=Treatment,
                 y=MeanProtein,
                 color=Organ),
             position = position_dodge(width=0.2))+
  labs(x= "Season of Last Burn",
       y= "Mean Protein (%)")
#Stats 
pmod<- lm(Protein~Treatment*Organ, cn)
anova(pmod) %>% pander::pander("Anova table")
TukeyHSD(aov(pmod), which= "Treatment") %>% pander::pander("Tukey test")
```

# Script
```{r script, ref.label=knitr::all_labels(), echo=TRUE, eval=FALSE}
```
